\Title

### 2021-08-27 latex/paper/main.tex
Second pass of the first part of the proof of Theorem 4

%/========== Proof ==========/%
\section{Proof of Theorem \ref{thm:acrobot-energy-stabilization}}\label{sec:proof}
In this section we prove Theorem \ref{thm:acrobot-energy-stabilization},
which claims the acrobot gains/loses energy on \(\mathcal{O}_1\) and
\(\mathcal{O}_2(\rho)\).
These are the sublevel sets \eqref{eqn:oscillation-domain} and
\eqref{eqn:o-rhobar} of the nominal pendulum's energy \(E(q_u,p_u)\) defined in
\eqref{eqn:nominal-energy}.

Here is fundamental idea behind this proof:
for each of \(\mathcal{O}_1\) and \(\mathcal{O}_2(\rho)\),
we can find a diffeomorphism \(T(q_u,p_u) = (r,\theta)\) into pseudo-polar
coordinates adapted to level sets of \(E(q_u,p_u)\);
that is, when \(I = 0\) in \eqref{eqn:acrobot-constraint}, the pseudo-radius
\(r\) chooses a level set of \(E\) and the pseudo-angle \(\theta\) chooses a
point on that level set.
If the constrained dynamics in \((r,\theta)\) coordinates flow in such a way
that \(r\) increases ``on average", then \(E\) must also increase ``on average",
and the original constrained dynamics \eqref{eqn:acrobot-constrained-dynamics} 
must be gaining energy on the domain of the diffeomorphism.

The following lemma provides a set of conditions on the \((r,\theta)\) dynamics
under which energy will increase on a domain \(D\); later in this section, we
will assign \(D \subset \mathcal{O}_1\) or \(D \subset \mathcal{O}_2(\rho)\),
find the diffeomorphism \(T\), and apply this lemma to prove the claims in
Theorem \ref{thm:acrobot-energy-stabilization}.

@@

    Since \((\partial g/\partial \hat{r})(\theta,r_0,0) \equiv 0\),
    the solution of \eqref{eqn:r1-hat-prime} can be obtained by
    quadrature, giving
    \(\hat{r}_1(\theta,r_0) = \int_0^\theta b(\sigma,r_0)d\sigma\).
%    \[
%        \hat{r}_1(\theta,r_0) = \int\limits_0^\theta b(\sigma,r_0)d\sigma
%        .
%    \]
%    with \(b(\theta,r_0) := (\partial g/\partial I)(\theta,r_0,0)\). 
    Recall that \(b(\theta,r_0)\) is \(2\pi\)-periodic in
    \(\theta\), so there is a well-defined Poincar\'{e} map describing
    how \(\hat{r}\) changes each time \(\theta\) increases by \(2\pi\).
    This map is 
    \begin{align}\label{eqn:poincare-map}
        P : \, ]\underline{r},\overline{r}[ &\to \mathbb{R}_{\geq 0} \\
        r_0 &\mapsto \hat{r}(2\pi,r_0,I)
        ,
    \end{align}
    which expands to \(P(r_0) = r_0 + IS(r_0) + R(2\pi,r_0,I)\).
    In other words, we can understand the evolution of \(\hat{r}\) by studying
    the dynamics of the discrete-time system
    \begin{equation}\label{eqn:r-discrete}
        r_{n+1} = P(r_n) =  r_n + IS(r_n) + R(2\pi,r_n,I)
        ,
    \end{equation}
    with initial condition \(r_0\).
%    Here, \(r_n\) corresponds to the distance along the line segment
%    \(L := ]\underline{r},\overline{r}[ \, \times \{0\}\)
%    when an orbit of \((r,\theta)\) hits \(L\) for the the \(n^\text{th}\) time,
%    assuming the acrobot was initialized at \((r,\theta) = (r_0,0)\).

    If at any point in discrete time \(r_n\) exits \(]\underline{r},\overline{r}[\), the orbit 
    in \((q_u,p_u)\) coordinates must have departed \(D\).
    Hence, proving the constrained dynamics gain energy on \(D\) is
    equivalent to finding \(N \in \mathbb{Z}_{> 0}\) such that 
    \(r_{n} \geq \overline{r}\) for all \(n \geq N\).
    Using a perturbation analysis (see Khalil 
    \cite[Theorem 10.1]{khalil_nonlinear}),
    we know there exists some \(I_2 \in \, ]0,I_1]\) and \(\kappa > 0\)
    such that \(\left|R(\theta,r,I)\right| \leq \kappa I^2\) for all 
    \(\theta \in [0,2\pi]\),
    \(r \in \, ]\underline{r},\overline{r}[\),
    and \(I \in [-I_2,I_2]\).
    Taking this fact alongside the assumption that \(S(r_0) \geq \epsilon\),
    we find a lower bound \(r_{n+1} \geq r_n + I\epsilon - \kappa I^2\).
    Let \(0 < I^\star < \min\{I_2, \epsilon/\kappa\}\).
    Then for each \(I \in ]0,I^\star]\), the value 
    \(\gamma := I\epsilon - \kappa I^2\) is strictly positive, so
    \(P(r_0) \geq r_0 + \gamma\).
    Therefore, \(r_{n} \geq r_0 + n\gamma\), 
    and all solutions of \eqref{eqn:r-discrete} eventually flow past
    \(\overline{r}\).

@@

\subsection{Energy Gain on \(\mathcal{O}_1\)}\label{sec:proof-o1}
To prove the first part of Theorem
\ref{thm:acrobot-energy-stabilization}, we use
\cite[Eqn. (9)]{dynamic_vhcs_stabilize_closed_orbits} to find 
pseudo-polar coordinates adapted to level sets of \(E\) on
\(\mathcal{O}_1 \backslash \{(0,0)\}\).
The resulting coordinate transformation is
\begin{align}
    \label{eqn:r-osc}
    r &= \arccos\left(\cos(q_u) - \frac{p_u^2}{30m^2gl^3}\right)
    , \\
    \label{eqn:theta-osc}
    \theta &=
    \left.\arctan_2\left(-\sign{p_u}\sqrt{1-\frac{q_u^2}{r^2}},\frac{q_u}{r}\right)
    \right|_{r = r(q_u,p_u)}
    ,
\end{align}
where \((r,\theta) \in \, ]0,\pi[ \times \Sone\).
One can verify that the inverse to this transformation is
\begin{align*}
    q_u &= rc_\theta
    , \\
    p_u &= -\sign{s_\theta}
    \sqrt{30m^2gl^3\cos(rc_\theta - c_r)}
    ,
\end{align*}
where for shorthand we write \(c_\theta := \cos(\theta)\), \(s_\theta :=
\sin(\theta)\), \(c_r := \cos(r)\), and \(s_r := \sin(r)\).
We now verify these pseudo-polar coordinates satisfy the requirements of 
Lemma \ref{lemma:energy-regulation-D}.
We omit the expressions for \(f_r(\theta,r,I)\) and \(f_\theta(\theta,r,I)\), as
they are too long for print.

It is straightforward to show that
\(f_r(\theta,r,0) \equiv 0\)
and
\begin{equation} \label{eqn:theta-dot-nom-osc}
    f_\theta(\theta,r,0) = \sqrt{\frac{6g}{5l}} 
        \sqrt{\frac{\cos(r c_\theta) - c_r}
            {r^2 s_\theta^2}}
    .
\end{equation}
Note that ~\eqref{eqn:theta-dot-nom-osc} has removable singularities at
\(\theta \in \{0,\pi\}\).
Taking the limit as \(\theta\) approaches these points yields
\begin{equation}\label{eqn:theta-dot-limit}
    \lim\limits_{\theta \to 0}f_\theta(\theta,r,0)
    = \lim\limits_{\theta \to \pi} f_\theta(\theta,r,0)
    = \sqrt{\frac{6g s_r}{10lr}}
    , 
\end{equation}
which is smooth and well-defined for all
\(r \in \, ]0,\pi[\).
From \eqref{eqn:theta-dot-nom-osc}--\eqref{eqn:theta-dot-limit},
one can verify that \(f_\theta(\theta,r,0) > 0\) as required.
Defining \(g(\theta,r,I)\) as in Lemma \ref{lemma:energy-regulation-D}, one can
also verify that \((\partial g/\partial r)(\theta,r,0) \equiv 0\).

Symbolic computations reveal that
\(b(\theta,r_0) = K a(\theta,r_0)\),
where
\begin{align*}
    K &:= \frac{\bar{q}_a \sqrt{30m^2g l^3}}{15}
    , \\
    a(\theta,r_0) &:= \frac{
        r_0 |s_\theta| \left(
        5 c_{r_0} \cos(r_0 c_\theta) - 8 \cos(r_0c_\theta)^2 + 3
    \right)
    }{
    s_{r_0}\sqrt{\cos(r_0c_\theta) - c_{r_0}}
    }
    .
\end{align*}
Notice that \(K\) is a positive constant which depends only on
\(m\), \(g\), \(l\), and \(\bar{q}_a\), while \(a(\theta,r_0)\) is adimensional
and \(2\pi\)-periodic in \(\theta\).
This means the integral of \(a(\theta,r_0)\) can be computed once for all
acrobots; 
hence, there exists \(\epsilon > 0\) such that 
\(S(r_0) := \int_0^{2\pi} b(\sigma,r_0)d\sigma \geq \epsilon\)
if and only if there exists \(\delta > 0\) such that 
\(Q(r_0) := \int_0^{2\pi} a(\sigma,r_0) d\sigma \geq \delta\)
for every \(r_0 \in ]0,\pi[\).

@@

Thankfully, we can still use Lemma \ref{lemma:energy-regulation-D} on an
appropriate subset of \(\mathcal{O}_1\).
To find this subset, first recall the discrete-time pseudo-radius dynamics
\eqref{eqn:r-discrete} from the proof of Lemma \ref{lemma:energy-regulation-D}.
Linearizing \eqref{eqn:r-discrete} at \(r_0 = 0\) yields
\[
    P^\prime(0) = 1 + I K Q^\prime(0) + R^\prime(2\pi,0,I)
    ,
\]
where the prime now denotes differentiation with respect to \(r\).
One can show that
\begin{align*}
    Q^\prime(0) = \int \limits_0^\pi 
    \lim \limits_{r_0 \to 0^+}\pdiff{a}{r}(\sigma,r_0) d\sigma
    = \frac{\pi}{2\sqrt{2}} > 0
    .
\end{align*}
Since \(R(2\pi,0,I)\) is \(O(I^2)\), so is
\(R^\prime(2\pi,0,I)\).
Hence, it can be written in the form
\(R^\prime(2\pi,0,I) = I^2 \tilde{R}(I)\) where \(\tilde{R}(I)\) is smooth and
zero at \(I = 0\).
Thus, there exists \(I_0 > 0\) such that 
\(I K Q^\prime(0) + I^2 \tilde{R}(I) > 0\) for all \(I \in \, ]0,I_0]\),
which implies \(P^\prime(0) > 1\).
The origin is now a repeller, which means there exists some
(unknown) \(\mu > 0\) where \(]0,\mu]\) is negatively
invariant for \eqref{eqn:r-discrete}, and all solutions in this
interval flow through \(r = \mu\).
If \(I < 0\), we have \(P^\prime(0) < 1\) and the origin is an
asymptotically stable equilibrium for \eqref{eqn:r-discrete} instead.

>>>>

%/========== Proof ==========/%
\section{Proof of Theorem \ref{thm:acrobot-energy-stabilization}}\label{sec:proof}
In this section we prove Theorem \ref{thm:acrobot-energy-stabilization},
which claims the acrobot constrained by the VNHC \eqref{eqn:acrobot-constraint}
will gain or lose energy on \(\mathcal{O}_1\) and \(\mathcal{O}_2(\rho)\) for
``small enough" values of \(I\).
%These are the sublevel sets \eqref{eqn:oscillation-domain} and
%\eqref{eqn:o-rhobar} of the nominal pendulum's energy \(E(q_u,p_u)\) defined in
%\eqref{eqn:nominal-energy}.

Here is intuition behind this proof:
for each of \(\mathcal{O}_1\) and \(\mathcal{O}_2(\rho)\),
we can find a diffeomorphism from \((q_u,p_u)\) into pseudo-polar
coordinates \((r,\theta)\) adapted to level sets of \(E(q_u,p_u)\);
that is, when \(I = 0\) in \eqref{eqn:acrobot-constraint}, the pseudo-radius
\(r\) chooses a level set of \(E\) and the pseudo-angle \(\theta\) chooses a
point on that level set.
We will show that \(r(t)\) increases ``on average", 
which implies \(E\) also increases ``on average",
and that the original constrained dynamics \eqref{eqn:acrobot-constrained-dynamics} 
are gaining energy on the domain of the diffeomorphism.

The following lemma provides a set of conditions on the \((r,\theta)\)-dynamics
under which energy will increase on a domain \(D \subset \SxR\); later in this section, we
will assign \(D \subset \mathcal{O}_1\) or \(D \subset \mathcal{O}_2(\rho)\)
and apply this lemma to prove the claims in Theorem
\ref{thm:acrobot-energy-stabilization}.

@@

    Since \((\partial g/\partial \hat{r})(\theta,r_0,0) \equiv 0\),
    the solution of \eqref{eqn:r1-hat-prime} can be obtained by
    quadrature, giving
    \(\hat{r}_1(\theta,r_0) = \int_0^\theta b(\sigma,r_0)d\sigma\).
%    \[
%        \hat{r}_1(\theta,r_0) = \int\limits_0^\theta b(\sigma,r_0)d\sigma
%        .
%    \]
%    with \(b(\theta,r_0) := (\partial g/\partial I)(\theta,r_0,0)\). 
    Recall that \(b(\theta,r_0)\) is \(2\pi\)-periodic in
    \(\theta\), so there is a well-defined Poincar\'{e} map describing
    how \(\hat{r}\) changes each time \(\theta\) increases by \(2\pi\).
    This map is 
    \begin{align}\label{eqn:poincare-map}
        P : \, ]\underline{r},\overline{r}[ &\to \mathbb{R}_{\geq 0} \\
        r_0 &\mapsto \hat{r}(2\pi,r_0,I) \nonumber
        ,
    \end{align}
    which expands to \(P(r_0) = r_0 + IS(r_0) + R(2\pi,r_0,I)\).
    In other words, we can understand the evolution of \(\hat{r}\) by studying
    the dynamics of the discrete-time system
    \begin{equation}\label{eqn:r-discrete}
        r_{n+1} = P(r_n) =  r_n + IS(r_n) + R(2\pi,r_n,I)
        ,
    \end{equation}
    with initial condition \(r_0\).
%    Here, \(r_n\) corresponds to the distance along the line segment
%    \(L := ]\underline{r},\overline{r}[ \, \times \{0\}\)
%    when an orbit of \((r,\theta)\) hits \(L\) for the the \(n^\text{th}\) time,
%    assuming the acrobot was initialized at \((r,\theta) = (r_0,0)\).

    If \(r_{n+1}\in \, ]\underline{r},\overline{r}[\), the orbit 
    \((q_u(t),p_u(t))\) starting at \(T^{-1}(r_n,0)\) remained in \(D\) as
    \(\theta\) increased by \(2\pi\) and
    landed at \(T^{-1}(r_{n+1},2\pi)\); otherwise, this orbit exited
    \(D\)  before \(\theta\) reached \(2\pi\).
    Hence, proving the constrained dynamics gain energy on \(D\) is
    equivalent to finding \(N \in \mathbb{Z}_{> 0}\) such that 
    \(r_{n} \geq \overline{r}\) for all \(n \geq N\).
    Using a perturbation analysis (see Khalil 
    \cite[Theorem 10.1]{khalil_nonlinear}),
    we know there exists some \(I_2 \in \, ]0,I_1]\) and \(\kappa > 0\)
    such that \(\left|R(\theta,r,I)\right| \leq \kappa I^2\) for all 
    \(\theta \in [0,2\pi]\),
    \(r \in \, ]\underline{r},\overline{r}[\),
    and \(I \in [-I_2,I_2]\).
    Taking this fact alongside the assumption that \(S(r_0) \geq \epsilon\),
    we find a lower bound \(r_{n+1} \geq r_n + I\epsilon - \kappa I^2\).
    Let \(0 < I^\star < \min\{I_2, \epsilon/\kappa\}\).
    Then for each \(I \in \, ]0,I^\star]\), the value 
    \(\gamma := I\epsilon - \kappa I^2\) is strictly positive, so
    \(P(r_0) \geq r_0 + \gamma\).
    Therefore, \(r_{n} \geq r_0 + n\gamma\), 
    and all solutions of \eqref{eqn:r-discrete} eventually flow past
    \(\overline{r}\).

@@

\subsection{Energy Gain on \(\mathcal{O}_1\)}\label{sec:proof-o1}
To prove the first part of Theorem
\ref{thm:acrobot-energy-stabilization}, we apply
\cite[Eqn. (9)]{dynamic_vhcs_stabilize_closed_orbits} to find 
pseudo-polar coordinates adapted to level sets of \(E\) on
\(\mathcal{O}_1 \backslash \{(0,0)\}\).
The resulting coordinate transformation is
\begin{align}
    \label{eqn:r-osc}
    r &= \arccos\left(\cos(q_u) - \frac{p_u^2}{30m^2gl^3}\right)
    , \\
    \label{eqn:theta-osc}
    \theta &=
    \left.\arctan_2\left(-\sign{p_u}\sqrt{1-\frac{q_u^2}{r^2}},\frac{q_u}{r}\right)
    \right|_{r = r(q_u,p_u)}
    ,
\end{align}
where \((r,\theta) \in \, ]0,\pi[ \times \Sone\).
One can verify that the inverse to this transformation is
\begin{align*}
    q_u &= rc_\theta
    , \\
    p_u &= -\sign{s_\theta}
    \sqrt{30m^2gl^3\cos(rc_\theta - c_r)}
    ,
\end{align*}
where for shorthand we write \(c_\theta := \cos(\theta)\), \(s_\theta :=
\sin(\theta)\), \(c_r := \cos(r)\), and \(s_r := \sin(r)\).
We now verify these pseudo-polar coordinates satisfy the requirements of 
Lemma \ref{lemma:energy-regulation-D}.
Note: we omit the expressions for \(f_r(\theta,r,I)\) and
\(f_\theta(\theta,r,I)\) as they are too long for print.

It is straightforward to show that
\(f_r(\theta,r,0) \equiv 0\)
and
\begin{equation} \label{eqn:theta-dot-nom-osc}
    f_\theta(\theta,r,0) = \sqrt{\frac{6g}{5l}} 
        \sqrt{\frac{\cos(r c_\theta) - c_r}
            {r^2 s_\theta^2}}
    .
\end{equation}
Observe that ~\eqref{eqn:theta-dot-nom-osc} has removable singularities at
\(\theta \in \{0,\pi\}\).
Taking the limit as \(\theta\) approaches these points yields
\begin{equation}\label{eqn:theta-dot-limit}
    \lim\limits_{\theta \to 0}f_\theta(\theta,r,0)
    = \lim\limits_{\theta \to \pi} f_\theta(\theta,r,0)
    = \sqrt{\frac{6g s_r}{10lr}}
    , 
\end{equation}
which is smooth and well-defined for all
\(r \in \, ]0,\pi[\).
From \eqref{eqn:theta-dot-nom-osc}--\eqref{eqn:theta-dot-limit},
one can verify that \(f_\theta(\theta,r,0) > 0\) as required.
Defining \(g(\theta,r,I)\) as in Lemma \ref{lemma:energy-regulation-D}, one can
also verify that \((\partial g/\partial r)(\theta,r,0) \equiv 0\).

Symbolic computations reveal that
\(b(\theta,r_0) = K a(\theta,r_0)\),
where
\begin{align*}
    K &:= \frac{\bar{q}_a \sqrt{30m^2g l^3}}{15}
    , \\
    a(\theta,r_0) &:= \frac{
        r_0 |s_\theta| \left(
        5 c_{r_0} \cos(r_0 c_\theta) - 8 \cos(r_0c_\theta)^2 + 3
    \right)
    }{
    s_{r_0}\sqrt{\cos(r_0c_\theta) - c_{r_0}}
    }
    .
\end{align*}
Notice that \(K\) is a positive constant which depends only on
\(m\), \(g\), \(l\), and \(\bar{q}_a\), while \(a(\theta,r_0)\) is adimensional
and \(2\pi\)-periodic in \(\theta\).
This means the integral of \(a(\theta,r_0)\) can be computed once for all
acrobots; 
hence, there exists \(\epsilon > 0\) such that 
\(S(r_0) := \int_0^{2\pi} b(\sigma,r_0)d\sigma \geq \epsilon\)
if and only if there exists \(\delta > 0\) such that 
\(Q(r_0) := \int_0^{2\pi} a(\sigma,r_0) d\sigma \geq \delta\)
for every \(r_0 \in \,]0,\pi[\).

@@

Thankfully, we can still use Lemma \ref{lemma:energy-regulation-D} on an
appropriate subset of \(\mathcal{O}_1\).
To find this subset, first recall the discrete-time pseudo-radius dynamics
\eqref{eqn:r-discrete} from the proof of Lemma \ref{lemma:energy-regulation-D}.
Linearizing \eqref{eqn:r-discrete} at \(r_0 = 0\) yields
\[
    P^\prime(0) = 1 + I K Q^\prime(0) + R^\prime(2\pi,0,I)
    ,
\]
where the prime now denotes differentiation with respect to \(r\).
One can show that
\begin{align*}
    Q^\prime(0) = \int \limits_0^\pi 
    \lim \limits_{r_0 \to 0^+}\pdiff{a}{r}(\sigma,r_0) d\sigma
    = \frac{\pi}{2\sqrt{2}} > 0
    .
\end{align*}
Since \(R(2\pi,0,I)\) is \(O(I^2)\), so is
\(R^\prime(2\pi,0,I)\).
Hence, it can be written in the form
\(R^\prime(2\pi,0,I) = I^2 \tilde{R}(I)\) where \(\tilde{R}(I)\) is smooth and
zero at \(I = 0\).
Thus, there exists \(I_0 > 0\) such that 
\(I K Q^\prime(0) + I^2 \tilde{R}(I) > 0\) for all \(I \in \, ]0,I_0]\),
which implies \(P^\prime(0) > 1\) for all \(I \in \, ]0,I_0]\).
This means the origin is a repeller, so there exists some
(unknown) \(\mu > 0\) where \(]0,\mu]\) is negatively
invariant for \eqref{eqn:r-discrete}, and all solutions in this
interval flow through \(r = \mu\).
If \(I < 0\), we have \(P^\prime(0) < 1\) and the origin is an
asymptotically stable equilibrium for \eqref{eqn:r-discrete}.
